Linear instability underlying synchronous waving of marine grass in flow

### 2014-09-07 Paper/grass.tex

\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{\text{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami }}
% \newcommand{\shreyas}[1]{{\bf (#1)}}
\newcommand{\shreyas}[1]{}
\definecolor{tableShade}{gray}{0.8}

@@

Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in Figure \ref{Re_vs_delta}. 
The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.

A threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), naturally emerges from the solution of \eqref{Orr-somerfield}. 
The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
The threshold Reynolds number increases with the grass density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
Previous linear stability calculations undertaken for terrestrial grass do not exhibit a threshold flow condition either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of the over-simplifying assumptions such as a tanh- or piece-wise linear mean velocity profile\cite{Raupach96,Delangre06}.
Regardless of the assumptions, the introduction of vegetation drag damped the linear modes.

@@

\begin{figure}
\includegraphics[]{Asymptotic_noshear}
\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density. Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached. Mode 2 shapes appear self-similar in shape as $\delta\to 0$, and are compared to each other on a rescaled coordinate in the inset. Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. The limit is not yet reached for the case $\delta/H = 0.01$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
\label{Asymptotic_mode}
\end{figure}

>>>>

\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{\text{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami}}
% \newcommand{\shreyas}[1]{{\bf (#1)}}
\newcommand{\shreyas}[1]{}
\definecolor{tableShade}{gray}{0.8}

@@

%<<<<<<< HEAD
% A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}. 
% The primary reason for the existence of critical $\Rey$ is the presence of dissipation of flow instability 
% due to grass drag. Indeed a calculation undertaken for terrestrial canopy shows the suppression of growth rate of instability
% due to grass drag\cite{Delangre06}. Absence of threshold in the previous analysis can be attributed to the assumptions such 
% as approximationn of mean flow with the free shear layer profile and the exclusion of grass drag in the stability analysis 
% \cite{Raupach96,Delangre06}.
% The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, 
% on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the 
% threshold\cite{Ghisal02}. The good agreement between predicted and estimated threshold suggest that the grass drag plays a 
% crucial role in setting up the flow instability.
% 
% Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in 
% Figure \ref{Re_vs_delta}. The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}\shreyas{cite someone}. The deviation of predicted frequency from the observed may be attributed to the various 
% simplification we have made in our model. In a real canopy, the drag coefficients are known to
% vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}. The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}. Although these variation
% are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the
% observed frequency and the predicted frequency.
% =======
Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, 
is also shown in Figure \ref{Re_vs_delta}. 
The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami and frequency of 
vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.

A threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), naturally emerges from the 
solution of \eqref{Orr-somerfield}. 
The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
The threshold Reynolds number increases with the grass density, indicating a competition between the 
destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
Previous linear stability calculations undertaken for terrestrial grass do not exhibit a threshold flow condition 
either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of the over-simplifying 
assumptions such as a tanh- or piece-wise linear mean velocity profile\cite{Raupach96,Delangre06}.
Regardless of the assumptions, the introduction of vegetation drag damped the linear modes.
%>>>>>>> bd11ad10d1dc4414f38f11c1e91e5e1d200223a2

@@

\begin{figure}
\includegraphics[]{Asymptotic_noshear}
\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness 
for representative values of grass density. Mode 1 is shown in solid and Mode 2 is shown in dashed. 
The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) 
but $\Rey/\Ndg = O(1)$. The mode shapes approach each other for these small values of $\delta/H$ indicating that 
the large grass density asymptote is reached. Mode 2 shapes appear self-similar in shape as $\delta\to 0$, 
and are compared to each other on a rescaled coordinate in the inset. Inset shows $|\phi|$ for 
Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit 
has been reached. The limit is not yet reached for the case $\delta/H = 0.05$ due to the influence of bottom boundary; 
the grass height in this case is comparable to the boundary layer thickness.}
\label{Asymptotic_mode}
\end{figure}
